Assessment and implementation of resource road simulation methods for projection of cumulative effects of disturbance on wildlife

### 2023-01-17 analysis/paper/paper.Rmd
minor edits, meaningful fig chunk labels and dynamic links

The impacts of roads on wildlife are complex and varied. Mortality risk and stress on or near roads can be elevated due to vehicle collisions, hunting pressure, predation, noise and other stressors, but roads can also provide food, facilitate movement, and provide refuge [@hill_review_2021;@trombulak_review_2000;@fahrig_effects_2009].    
In Canadian forests, there is ample evidence of linear features attracting predators such as bears (*Ursus americanus, Ursus arctos*) and wolves (*Canis lupus*) seeking food or easy movement, while repelling prey such as woodland caribou (*Rangifer tarandus caribou*) and moose (*Alces americanus*) [@dickie_corridors_2020;@polfus_identifying_2011;@mumma_predation_2018;@beauchesne_disentangling_2013;@leblond_avoidance_2013;@laurian_interactions_2012]; there is also evidence of predators avoiding features heavily used by humans [@muhly_human_2011;@whittington_towns_2022; @proctor_effects_2020]. Changes in mortality risks, habitat quality, and connectivity associated with anthropogenic disturbance can in turn alter use and movement patterns [@tucker_moving_2018; @apps_factors_2006; @proctor_effects_2020], demography and population viability [@johnson_science_2020; @decesare_linking_2014; @barrientos_lost_2021;@apps_spatial_2013;@pinard_calving_2012;@fahrig_effects_2009; @lochhead_linking_2022; @apps_spatial_2013; @proctor_effects_2020]. Roads can also facilitate spread of invasive species [@riitters_landscape_2017], and enable other development activities [@johnson_growth-inducing_2020].

@@

<!-- TO DO: dynamic links to figs and tables. -->
We chose the Revelstoke timber supply area (TSA 27), in British Columbia, Canada as an assessment area because forestry is the dominant resource extraction activity in this region and much of the road network was built to access to forest harvest cutblocks (Figure 1). This allows projection of the forest road network from historic forest harvest data. Approximately 55% of the area is non-forested land where road building is difficult or impossible (e.g. alpine, lakes, swamp, brush, rock) [@hachey_revelstoke_2010], so the area provides opportunities to observed the responses of road projection algorithms to barriers. Revelstoke is in the interior-wet belt of the province where forests are dominated by Englemann spruce (31%), Hemlock (23%), Cedar (22%), and Douglas-fir (18%) [@ministry_of_forests_revelstoke_2021]. The area is home to wildlife species that are sensitive to roads including grizzly bears and Southern Mountain woodland caribou [@hachey_revelstoke_2010]. Eight herds of Southern Mountain caribou have been extirpated or likely extirpated, and remaining herds are considered Endangered by COSEWIC and listed as threatened under the Canadian Species at Risk Act [@palm_long_2020]. Mountain caribou avoid roads leading to a loss of available habitat [@polfus_identifying_2011; @apps_factors_2006], suffer increased mortality near roads [@apps_spatial_2013], and decline in abundance as the footprint of roads and cutblocks increases [@lochhead_linking_2022]. Roads also alter grizzly habitat use, home range selection, movements, population fragmentation, survival, reproductive rates, population density, trends, and conservation status [@proctor_effects_2020].

@@

```{r fig1, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel.", cache=TRUE}
tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
tsaCost <- terra::rast(here(data_path_drvd, "input_cost.tif"))

@@

To spatially define the cost of building and traveling on a forestry road we developed a simple model informed by slope and major barriers (Table 1). Following the Interior Appraisal Manual of BC, we assume the base cost per kilometer of road is a linear function of slope [@government_of_british_columbia_interior_2021], and that road construction across major water bodies is prohibitively expensive (Table 1). Existing roads are assigned a cost of 0. The cost raster consists of 1703936 pixels with a resolution of 1 ha.

Table 1. A simple model of road construction costs.
```{r}
tbl <- "Cost Type|Value ($ per km)|Source
Slope|16178 + 504*slope|Interior appraisal manual
Major lakes|65000|Major water bodies of BC
Existing roads|0|Digital road atlas and forest tenure roads for BC
"
df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,]))
df <- df[-1,] # remove first row
row.names(df)<-NULL
knitr::kable(df)
```

@@

Simple methods for projecting low-cost routes from an existing road network to target locations (i.e. harvest cutblocks) include simple least cost paths (LCP), dynamic least cost paths (DLCP) and least cost path minimum spanning trees (MST) \@ref(fig:fig2)). We use Dijkstra's shortest path algorithm [-@dijkstra_note_1959] to determine the least cost path between a target location (e.g. centroid of a cutblock) and the existing road network. Least cost paths to each cutblock is determined independently in simple method (LCP). The dynamic least cost paths (DLCP) method evaluates the path to each cutblock in sequence so that earlier paths are considered in the cost for later paths. Sequence is determined by a heuristic such as the distance to the nearest existing road in these examples. The minimum spanning tree (MST) method calculates the minimum cost network to connecting all cutblocks to the road network using Kruskal's algorithm [-@kruskal_shortest_1956]. See Appendix A for a more detailed explanation of these methods. We focus our analysis on the DLCP and MST approaches which yield more realistic road networks (Figure \@ref(fig:fig2)). We compared these methods to an alternative iterative least cost path method developed by Hardy [-@hardy_forest_2019].

@@

The DLCP and MST methods connect target or landing locations within each cutblock to the existing road network, so an important first step is to select one or more landing locations in each block. We investigated the sensitivity of results to the pattern and density of landing locations by comparing results derived from the centroid of each cutblock to regular or randomly selected landing densities of 1 (low) and 10 (high) points per km^2^ (Figure \@ref(fig:fig2)). We used default settings in the Hardy QGIS plugin except for kidding distance which was set to 200m. A skidding distance of 200 m corresponds to \~ 9 pts/km^2^. We compared two resolutions of the cost surface by aggregating the 1 ha resolution cost raster to a coarser 100 ha (1000 by 1000 m) resolution. In total, we compared 12 road projection method variants to a cutblocks only scenario with no projected roads (Table 2).

Table 2: Road projection method variants. 
```{r}
tbl2 <- "Projection method|Sample type|Sample density (pts/km^2^)|Resolutions
MST|Regular|1|1 ha and 100 ha
MST|Regular|10|1 ha and 100 ha
MST|Random|1|1 ha and 100 ha
MST|Random|10|1 ha and 100 ha
MST|Centroid|N/A|1 ha and 100 ha
DLCP|Regular|10|1 ha 
Hardy QGIS|N/A|~9|1 ha
Cutblocks only|N/A|N/A|1 ha

"
df2 <- read.delim(textConnection(tbl2),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df2) <- unname(as.list(df2[1,]))
df2 <- df2[-1,] # remove first row
row.names(df2)<-NULL
knitr::kable(df2)
```

@@

```{r fig2, fig.cap="We compared minimum spanning tree (MST), dynamic least cost paths (DLCP) and least cost path (LCP) road network projection algorithms. Landing locations were selected as the centroid of the cutblock, or a random or regularly spaced sample of points inside each cutblock (Table 2). Target cutblocks are outlined in grey, existing roads are black, projected roads are yellow, observed roads are red. Cost (1 ha resolution) is indicated by grey shading."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.png"))
```

@@

For each projection method variant the aspatial performance varied across the five road network metrics and also varied depending on whether we consider scores within cutblocks or overall scales (Figure \@ref(fig:fig3)). All road network projection method variants accurately predict the forestry disturbance footprint, and slightly overproject the road disturbance footprint within cutblocks. The high density methods overproject the overall road disturbance footprint more than low density methods, and underproject overall distance to roads, but all method variants give reasonably good projections of overall distance to roads. All methods underproject road density, in part because within 1 ha pixels all the methods generate straight road segments. All the methods also underproject road presence, and the problem is more severe for the low density method variants. In general, performance differences among low and high density method variants are larger than differences among projection algorithms. With the coarser resolution cost raster road projections had mean road presence that was closer to the observed, while road density was further from the observed mean and distance to the nearest road was further from the observed mean inside cutblocks but closer for the landscape overall (Figure A-\@ref(fig:figA1)).

@@

```{r fig3, fig.height=3, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection method variants with a 1 ha resolution cost raster across the entire Revelstoke Timber Supply Area (“overall”) and within cutblocks (“cutblocks”). Values greater than 0 indicate overprojection of the metric and negative values indicate an underprojection. The cutblocks only method is not defined for the road density, road presence or distance to road characteristics."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA NA klementQGIS",
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

Difference maps highlight several potential causes of discrepancy between our projections and the observed road network. In most areas the projections remained within a reasonable vicinity of the observed roads and also realistically remained within valley bottoms without cutting across natural barriers such as mountains (Figure \@ref(fig:fig4)a). However, the cost raster is not sufficiently details to reflect costs of crossing rivers, streams or other wet areas in valley bottoms so the projections cross valleys more than the observed roads which often run parallel to one another on either side of valleys. There were also instances where cutblocks included in the data set appear to be misidentified and are not accessed by the observed road network (Figure \@ref(fig:fig4)b). In addition, the projected roads were not able to match the switchback pattern of observed roads in steep terrain (Figure \@ref(fig:fig4)c). All these discrepancies could be addressed with more accurate harvest and cost information.

```{r fig4, fig.height=9, fig.width=fig_widths["two"], fig.cap="Comparison of projection results for the minimum spanning tree projection with regular high density sampling to the observed road network. The primary map shows the observed distance to the nearest road minus the projected distance and the insets show close-ups of the road networks with satellite imagery to give more context. a) Projected and observed road networks follow similar patterns although projected roads cross valley bottoms more frequently. b) An example of a misidentified cutblock. c) Projected roads do not match observed switchback roads.  The satellite images are from Google (2022)."}

@@

```{r fig5, fig.height=4.5, fig.width=fig_widths["two"], fig.cap="Comparison of performance metrics for each projection method variant using the 1 ha resolution cost raster. F-measure is the harmonic mean of precision and sensitivity, precision is the proportion of predicted roads that are observed roads and sensitivity is the proportion of observed roads that were correctly predicted. Values closer to one indicate better performance."}

@@

As expected, none of the projection methods can accurately project road presence (i.e. the exact locations of roads) (F-Measure: `r perf_tbl_rng["Road presence"]`), and all of the methods provide accurate projections of the forestry disturbance footprint (F-Measure: `r perf_tbl_rng["Forestry disturbance\nfootprint"]`). All methods also provide reasonably good projections of the road disturbance footprint (F-Measure: `r perf_tbl_rng["Road disturbance\nfootprint"]`), though higher sampling densities yield fewer false negatives and more false positives. All projection methods yielded more accurate results than the 'cutblock only' scenario, indicating that any of these projection methods are better than none (Figure \@ref(fig:fig5)). The DLCP projection method had slightly lower performance than the MST projection method with the same sampling strategy. The DLCP method was faster to compute however, with a run time of `r dlcp_time` minutes compared to `r mst_time` minutes for the MST method.

>>>>

The impacts of roads on wildlife are complex and varied. Mortality risk and stress on or near roads can be elevated due to vehicle collisions, hunting pressure, predation, noise and other stressors, but roads can also provide food, facilitate movement, and provide refuge [@hill_review_2021;@trombulak_review_2000;@fahrig_effects_2009].    
In Canadian forests, there is ample evidence of linear features attracting predators such as bears (*Ursus americanus*, *Ursus arctos*) and wolves (*Canis lupus*) seeking food or easy movement, while repelling prey such as woodland caribou (*Rangifer tarandus caribou*) and moose (*Alces americanus*) [@dickie_corridors_2020;@polfus_identifying_2011;@mumma_predation_2018;@beauchesne_disentangling_2013;@leblond_avoidance_2013;@laurian_interactions_2012]; there is also evidence of predators avoiding features heavily used by humans [@muhly_human_2011;@whittington_towns_2022; @proctor_effects_2020]. Changes in mortality risks, habitat quality, and connectivity associated with anthropogenic disturbance can in turn alter use and movement patterns [@tucker_moving_2018; @apps_factors_2006; @proctor_effects_2020], demography and population viability [@johnson_science_2020; @decesare_linking_2014; @barrientos_lost_2021;@apps_spatial_2013;@pinard_calving_2012;@fahrig_effects_2009; @lochhead_linking_2022; @apps_spatial_2013; @proctor_effects_2020]. Roads can also facilitate spread of invasive species [@riitters_landscape_2017], and enable other development activities [@johnson_growth-inducing_2020].

@@

We chose the Revelstoke timber supply area (TSA 27), in British Columbia, Canada as an assessment area because forestry is the dominant resource extraction activity in this region and much of the road network was built to access to forest harvest cutblocks (Figure \@ref(fig:study-area)). This allows projection of the forest road network from historic forest harvest data. Approximately 55% of the area is non-forested land where road building is difficult or impossible (e.g. alpine, lakes, swamp, brush, rock) [@hachey_revelstoke_2010], so the area provides opportunities to observed the responses of road projection algorithms to barriers. Revelstoke is in the interior-wet belt of the province where forests are dominated by Englemann spruce (31%), Hemlock (23%), Cedar (22%), and Douglas-fir (18%) [@ministry_of_forests_revelstoke_2021]. The area is home to wildlife species that are sensitive to roads including grizzly bears and Southern Mountain woodland caribou [@hachey_revelstoke_2010]. Eight herds of Southern Mountain caribou have been extirpated or likely extirpated, and remaining herds are considered Endangered by COSEWIC and listed as threatened under the Canadian Species at Risk Act [@palm_long_2020]. Mountain caribou avoid roads leading to a loss of available habitat [@polfus_identifying_2011; @apps_factors_2006], suffer increased mortality near roads [@apps_spatial_2013], and decline in abundance as the footprint of roads and cutblocks increases [@lochhead_linking_2022]. Roads also alter grizzly habitat use, home range selection, movements, population fragmentation, survival, reproductive rates, population density, trends, and conservation status [@proctor_effects_2020].

@@

```{r study-area, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel.", cache=TRUE}
tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
tsaCost <- terra::rast(here(data_path_drvd, "input_cost.tif"))

@@

To spatially define the cost of building and traveling on a forestry road we developed a simple model informed by slope and major barriers (Table 1). Following the Interior Appraisal Manual of BC, we assume the base cost per kilometer of road is a linear function of slope [@government_of_british_columbia_interior_2021], and that road construction across major water bodies is prohibitively expensive (Table \@ref(tab:cost-tbl)). Existing roads are assigned a cost of 0. The cost raster consists of 1703936 pixels with a resolution of 1 ha.

```{r cost-tbl}
tbl <- "Cost Type|Value ($ per km)|Source
Slope|16178 + 504*slope|Interior appraisal manual
Major lakes|65000|Major water bodies of BC
Existing roads|0|Digital road atlas and forest tenure roads for BC
"
df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,]))
df <- df[-1,] # remove first row
row.names(df)<-NULL
knitr::kable(df, caption = "A simple model of road construction costs.")
```

@@

Simple methods for projecting low-cost routes from an existing road network to target locations (i.e. harvest cutblocks) include simple least cost paths (LCP), dynamic least cost paths (DLCP) and least cost path minimum spanning trees (MST) (Figure \@ref(fig:proj-meth-fig)). We use Dijkstra's shortest path algorithm [-@dijkstra_note_1959] to determine the least cost path between a target location (e.g. centroid of a cutblock) and the existing road network. Least cost paths to each cutblock is determined independently in simple method (LCP). The dynamic least cost paths (DLCP) method evaluates the path to each cutblock in sequence so that earlier paths are considered in the cost for later paths. Sequence is determined by a heuristic such as the distance to the nearest existing road in these examples. The minimum spanning tree (MST) method calculates the minimum cost network to connecting all cutblocks to the road network using Kruskal's algorithm [-@kruskal_shortest_1956]. See Appendix A for a more detailed explanation of these methods. We focus our analysis on the DLCP and MST approaches which yield more realistic road networks (Figure \@ref(fig:proj-meth-fig)). We compared these methods to an alternative iterative least cost path method developed by Hardy [-@hardy_forest_2019].

@@

The DLCP and MST methods connect target or landing locations within each cutblock to the existing road network, so an important first step is to select one or more landing locations in each block. We investigated the sensitivity of results to the pattern and density of landing locations by comparing results derived from the centroid of each cutblock to regular or randomly selected landing densities of 1 (low) and 10 (high) points per km^2^ (Figure \@ref(fig:proj-meth-fig)). We used default settings in the Hardy QGIS plugin except for kidding distance which was set to 200m. A skidding distance of 200 m corresponds to \~ 9 pts/km^2^. We compared two resolutions of the cost surface by aggregating the 1 ha resolution cost raster to a coarser 100 ha (1000 by 1000 m) resolution. In total, we compared 12 road projection method variants to a cutblocks only scenario with no projected roads (Table \@ref(tab:method-tbl)).

```{r method-tbl}
tbl2 <- "Projection method|Sample type|Sample density (pts/km^2^)|Resolutions
MST|Regular|1|1 ha and 100 ha
MST|Regular|10|1 ha and 100 ha
MST|Random|1|1 ha and 100 ha
MST|Random|10|1 ha and 100 ha
MST|Centroid|N/A|1 ha and 100 ha
DLCP|Regular|10|1 ha 
Hardy QGIS|N/A|~9|1 ha
Cutblocks only|N/A|N/A|1 ha

"
df2 <- read.delim(textConnection(tbl2),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df2) <- unname(as.list(df2[1,]))
df2 <- df2[-1,] # remove first row
row.names(df2)<-NULL
knitr::kable(df2, caption = "Road projection method variants.")
```

@@

```{r proj-meth-fig, fig.cap="We compared minimum spanning tree (MST), dynamic least cost paths (DLCP) and least cost path (LCP) road network projection algorithms. Landing locations were selected as the centroid of the cutblock, or a random or regularly spaced sample of points inside each cutblock (Table 2). Target cutblocks are outlined in grey, existing roads are black, projected roads are yellow, observed roads are red. Cost (1 ha resolution) is indicated by grey shading."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.png"))
```

@@

For each projection method variant the aspatial performance varied across the five road network metrics and also varied depending on whether we consider scores within cutblocks or overall scales (Figure \@ref(fig:aspat-perf-fig)). All road network projection method variants accurately predict the forestry disturbance footprint, and slightly overproject the road disturbance footprint within cutblocks. The high density methods overproject the overall road disturbance footprint more than low density methods, and underproject overall distance to roads, but all method variants give reasonably good projections of overall distance to roads. All methods underproject road density, in part because within 1 ha pixels all the methods generate straight road segments. All the methods also underproject road presence, and the problem is more severe for the low density method variants. In general, performance differences among low and high density method variants are larger than differences among projection algorithms. With the coarser resolution cost raster road projections had mean road presence that was closer to the observed, while road density was further from the observed mean and distance to the nearest road was further from the observed mean inside cutblocks but closer for the landscape overall (Figure A-\@ref(fig:figA1)).

@@

```{r aspat-perf-fig, fig.height=3, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection method variants with a 1 ha resolution cost raster across the entire Revelstoke Timber Supply Area (“overall”) and within cutblocks (“cutblocks”). Values greater than 0 indicate overprojection of the metric and negative values indicate an underprojection. The cutblocks only method is not defined for the road density, road presence or distance to road characteristics."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA NA klementQGIS",
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

Difference maps highlight several potential causes of discrepancy between our projections and the observed road network. In most areas the projections remained within a reasonable vicinity of the observed roads and also realistically remained within valley bottoms without cutting across natural barriers such as mountains (Figure \@ref(fig:proj-map-fig)a). However, the cost raster is not sufficiently details to reflect costs of crossing rivers, streams or other wet areas in valley bottoms so the projections cross valleys more than the observed roads which often run parallel to one another on either side of valleys. There were also instances where cutblocks included in the data set appear to be misidentified and are not accessed by the observed road network (Figure \@ref(fig:proj-map-fig)b). In addition, the projected roads were not able to match the switchback pattern of observed roads in steep terrain (Figure \@ref(fig:proj-map-fig)c). All these discrepancies could be addressed with more accurate harvest and cost information.

```{r proj-map-fig, fig.height=9, fig.width=fig_widths["two"], fig.cap="Comparison of projection results for the minimum spanning tree projection with regular high density sampling to the observed road network. The primary map shows the observed distance to the nearest road minus the projected distance and the insets show close-ups of the road networks with satellite imagery to give more context. a) Projected and observed road networks follow similar patterns although projected roads cross valley bottoms more frequently. b) An example of a misidentified cutblock. c) Projected roads do not match observed switchback roads.  The satellite images are from Google (2022)."}

@@

```{r spat-perf-fig, fig.height=4.5, fig.width=fig_widths["two"], fig.cap="Comparison of performance metrics for each projection method variant using the 1 ha resolution cost raster. F-measure is the harmonic mean of precision and sensitivity, precision is the proportion of predicted roads that are observed roads and sensitivity is the proportion of observed roads that were correctly predicted. Values closer to one indicate better performance."}

@@

As expected, none of the projection methods can accurately project road presence (i.e. the exact locations of roads) (F-Measure: `r perf_tbl_rng["Road presence"]`), and all of the methods provide accurate projections of the forestry disturbance footprint (F-Measure: `r perf_tbl_rng["Forestry disturbance\nfootprint"]`). All methods also provide reasonably good projections of the road disturbance footprint (F-Measure: `r perf_tbl_rng["Road disturbance\nfootprint"]`), though higher sampling densities yield fewer false negatives and more false positives. All projection methods yielded more accurate results than the 'cutblock only' scenario, indicating that any of these projection methods are better than none (Figure \@ref(fig:spat-perf-fig)). The DLCP projection method had slightly lower performance than the MST projection method with the same sampling strategy. The DLCP method was faster to compute however, with a run time of `r dlcp_time` minutes compared to `r mst_time` minutes for the MST method.
